Robustly decorrelating errors with mixed quantum gates

### 2019-09-12 decorrelation/mixed_unitaries.tex
Finished up to section IVB

\newcommand{\actual}{\ensuremath{\tilde{\mathsf{G}}}}
\newcommand{\actmat}{\ensuremath{\tilde{\mathcal{G}}}}
\newcommand{\target}{\ensuremath{{\mathsf{G}}}}
\newcommand{\tarmat}{\ensuremath{{\mathcal{G}}}}
\newcommand{\error}{\ensuremath{{\mathsf{E}}}}
\newcommand{\errmat}{\ensuremath{{\mathcal{E}}}}
\newcommand{\genmat}{\ensuremath{{\mathcal{L}}}}
\newcommand{\vectorize}[1]{\ensuremath{\mathsf{vec}\left(#1\right)}}
\newcommand{\AGI}{\ensuremath{\epsilon_\mathcal{F}}}
\newcommand{\dnorm}{\ensuremath{\epsilon_\diamond}}
\newcommand{\identmat}{\ensuremath{\mathcal{I}}}
\newcommand{\ident}{\ensuremath{\mathsf{I}}}
\newcommand{\weight}{\ensuremath{w}}

\def\id{\mbox{\small 1} \!\! \mbox{1}}

@@

% ==============================================================================
% Section: Introduction
% ==============================================================================
\section{Introduction}
\label{sec:introduction}
\noindent The ultimate impact of a gate error on a quantum circuit depends strongly on both the magnitude and the nature of the error. Systematic, or \emph{coherent}, errors can arise from poorly calibrated controls or approximate gate compilations that induce repeatable, undesired unitary errors on the state of a quantum information processor (QIP). Errors of this type are correlated in time may add up constructively or destructively, depending on the circuit. They are are  computationally expensive to model and, because coherent errors may amplify when gates are repeated, it can be difficult to place tight analytic bounds on circuit performance \cite{Beale2018}. Contrast this against random, or \emph{stochastic}, errors which often result from high-frequency noise in the controls or the environment. Systems with stochastic errors can usually be accurately modeled by defining a rate of various discrete errors in the system, such as a bit flips or phase flips. These errors are significantly easier to simulate on a classical computer, and their impact on quantum circuits is much easier to estimate \cite{Beale2018}.

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{simple_example.pdf}
  \caption{A simple example of a mixed quantum gate. Using optimal control, two implementations of a $Z_\pi$ gate are designed to have equal and opposite coherent errors (if one implementation over-rotates by a small angle $\theta$, then the other \emph{under}-rotates by $\theta$). Each time the gate is used, one of these two implementations is chosen at random. The resulting effective quantum gate is equivalent to a perfect implementation of the gate followed by dephasing with associated probability $\order{\theta^2}$.}
  \label{fig:simple_example}
\end{figure}

Campbell \cite{Campbell2017,1811.08017} and Hastings \cite{Hastings2017} have shown that coherent noise can be strongly suppressed by probabilistically mixing several distinct implementations of the target quantum gates. They focus on errors arising from various gate compilations, such as the Solovey-Kitaev algorithm, for which any lingering approximation errors are generally coherent, even if the underlying native gates are perfect. Different approximate gate compilations of the same target unitary will almost certainly result in a different unitary errors. So by selecting from these various compilations at random, the resulting quantum channel becomes a mixture of unitaries \cite{DBLP:journals/corr/cs-CC-0012017}, which can have significantly less coherent error than any single compilation on its own. A very simple example of this reduction in coherent error by mixed quantum gates is illustrated in Fig.~\ref{fig:simple_example}.

In this article, we extend the work of Campbell and Hastings to numerically optimized quantum gates and show that the advantages of their approach can be made robust to drift in the gate implementations. We demonstrate that, depending on what properties of the resulting channel are desired, different numerical optimization targets may be preferred. We present an experimental implementation of single-qubit mixed unitary controls on a superconducting qubit testbed at Rigetti Computing. Using randomized benchmarking, we are able to show a marked improvement in error rates, as well as a reduced variance in circuit outcome probabilities, consistent with a significant reduction in the coherence of the error \cite{Ball2016}. We further apply our methods in simulation, constructing single- and two-qubit mixed unitary controls that are robust to drift and uncertainty in the control parameters.

% Reducing the coherence of a the quantum gate error

% 
%

% It has long been recognized, however, that one may inject additional decorrelating randomness

% By injecting additional decorrelating randomness into the system,

% By introducing additional decorrelating randomness, a number of techniques are able to convert coherent  errors into stochastic errors.

% By introducing intentional additional randomness to the 
% The gate-to-gate correlations that occur in the presence of coherent noise

% \note{don't define diamond norm here}
% The diamond distance can be used to bound the total variation distance (TVD) of a quantum circuit, but it is in general sensitive at first order to repeated application of a gate with coherent errors. For long circuits, this can add up extremely quickly, and Additionalf`while these errors can often be identified and reconstructed using various tomographic techniques, their impact on a given quantum  difficult to predict. Additionally, despite the relative ease of modeling stochastic errors, coherent errors are often much more likely to appear in QIPs.

%Drifting control parameters or environmental variables can easily have long correlation times that result in errors which are strongly coherent over the length of a quantum circuit.

@@

\subsection{Representing Errors in Quantum Gates}
\label{sec:rep_errors}
\noindent Suppose that one wishes to apply a unitary gate, $\target$, on some quantum device. Implementing this operation on a real system generally involves the application of a sequence of classical control fields to some set of qubits.  But fluctuations in the environment or imperfections in these controls can cause the state of the qubits to change in a way that is different from what was intended, \emph{i.e.}, there are errors in the gate. If the device is fairly stable with time and context\cite{Rudinger2019}, then we can accurately model the gate action using a completely positive, trace-preserving (CPTP) map, $\actual$ acting on the Hilbert space of the target qubits. This map can always be written as $\actual = \error\circ\target$, where $\error = \actual\circ\target^{-1}$ is the \emph{error map}, which is itself CPTP because $\target$ is unitary.

CPTP maps possess a number of useful representations, including Kraus operators\cite{1983}, Choi matrices\cite{Choi1975}, and Jamiolkowsi states\cite{yczkowski2004}. But for the purposes of this article, the \emph{process matrix} representation will be particularly convenient\cite{OBrien2004}, and we shall denote the process matrix associated with a given CPTP map $\target$ with its corresponding calligraphic character, $\tarmat$. For a $d$-dimensional quantum state, the process matrix is a $d^2\times d^2$ matrix. A key feature of process matrices is that they are composable and act through the usual matrix multiplication on the vectorized quantum state:
\begin{align}
	\vectorize{\actual(\rho)}
		&= \actmat\cdot\vectorize{\rho} \\
		&= \errmat\cdot\tarmat\cdot\vectorize{\rho}
\end{align}
The $\mathsf{vec}$ operation is typically performed using a basis of matrix units, for which $\vectorize{\rho}$ would be the column vector obtained by stacking the columns of $\rho$. In this work, however, we shall use a basis of Pauli matrices, defining $P = \left\{I, \sigma_x, \sigma_y, \sigma_z\right\}^{\otimes n}$ as the collection of all $4^n$ $n$-qubit Pauli operators (including the identity).  In this basis,
\begin{equation}
  \vectorize{\rho}_i = \expect{P_i} = \tr\!\left(P_i\; \rho \right),
\end{equation}
and
\begin{equation}
	(\actmat)_{ij} = \tr\!\left(P_i \;\actual\!\left(P_j\right) \right).
\end{equation}
Process matrices written in this Pauli basis are often referred to as \emph{Pauli transfer matrices} \cite{Chow2012}.  Error maps represented in this basis take the particularly nice form,
\begin{equation}\label{eq:process_matrix}
\errmat =
	\left(\begin{array}{c|cccc}
		1 &  & \vec{0}^T & \\ 
		\hline & &  &  \\
		\vec{m} &  & R &  \\
		 &  &  & 
	\end{array} 	
	\right)
\end{equation}
The top row of all trace-preserving (TP) error maps is fixed to $\{1,0,0,0,\cdots\}$ and the remainder of the first column, $\vec{m}$, describes any deviations from unitality, as might arise from amplitude damping \cite{preskill1997lecture}. If the error map is unitary, then the error is called \textit{coherent} and the unital submatrix $R$ is a rotation matrix.  Importantly, if $R$ is diagonal, then the error is Pauli stochastic, with each diagonal entry corresponding to the probability that its associated Pauli error occurs in each application of the gate.

In what follows it will be useful to define the \emph{error generator}, $\genmat$, associated with a faulty gate: 
\begin{align}
	\label{eq:generator}
	\errmat 
		&= \exp\left(\genmat\right) \\
		&= \identmat_d + \genmat + \frac{1}{2}\genmat^2 + \order{\genmat^3}.
\end{align}
If an implemented gate is relatively close to its target, then the error generator will be small under any of the usual matrix norms, and the Taylor expansion above may reliably be truncated at first or second order.

\subsection{Mixed Quantum Gates}
\label{sec:mqg}
\noindent Suppose that we have access to an ensemble of distinct implementations, 
$\{\actual_i\}$, of a target gate, $\target$.
Each time the gate is to be applied to the system, we randomly select an implementation from this ensemble such that the probability of drawing $\actual_i$ is $\weight_i$ (and we ensure that $\sum_i \weight_i=1$). This procedure is operationally indistinguishable from always applying the effective channel, 
\begin{align}
	\actual_{\rm eff} &= \sum_i \weight_i \actual_i = \left(\sum_i \weight_i \error_i \right) \circ\target
\end{align}
We call such randomized quantum operations \emph{mixed quantum gates} or MQGs. Error metrics for these MQGs can be computed in terms of their effective error map, 
\begin{equation}
	\label{eq:effective_error}
	\error_{\rm eff} = \sum_i \weight_i \error_i.
\end{equation}
Two important error metrics are the average gate infidelity (AGI), $\AGI$, and the diamond distance to the target (or simply, the \emph{diamond distance}), $\dnorm$ \cite{watrous2018theory}, defined as:
\begin{align}
	\label{eq:def_agi}
	&\AGI(\error) = \frac{d^2 - \tr{\errmat}}{d^2 + d},\\
	\label{eq:def_diamond}
	&\dnorm \left(\error\right)
		= \frac{1}{2} \sup_\rho \vert \vert (\ident_d\otimes \ident_d)(\rho) 
										  - (\error \otimes \ident_d)(\rho) \vert\vert_1,
\end{align}
where $d=2^n$ and $\ident_d$ is the $d$-dimensional identity operator.  In Eq.~\eqref{eq:def_agi}, we have written the AGI for an error map, $\error$, in terms of its associated Pauli transfer matrix, $\errmat$. If the error channel is purely stochastic, then $\dnorm(\error) = \AGI(\error)$, but if the error channel has a unitary component, then the diamond distance will generically be larger than the average gate infidelity \cite{1511.00727}. The diamond distance is subaddative \cite{watrous2018theory} under gate composition, so is particularly useful for constructing bounds on quantum circuit performance: the total variation distance between the outcome probabilities of a faulty and perfect quantum circuit is less than or equal to the sum of the diamond distances for gates that compose the circuit \cite{aharonov1998quantum}.

@@

Campbell \cite{Campbell2017} and Hastings \cite{1612.01011} independently considered these mixed channels using gates constructed with the Solovey-Kitaev algorithm, for which many approximate gate compilations are possible. Campbell showed that, if the error generators of some ensemble of gate compilations form a convex set containing the origin, then one can construct a MQG with quadratically suppressed diamond distance to the target. Explicitly, the weights are chosen such that the error generator is canceled to first order. Using (\ref{eq:effective_error}) and (\ref{eq:generator}), the effective error map for a MQG in terms of the component error generators is,
\begin{align}
	\label{eq:generators_and_effective_error}
	% \errmat_{\rm eff} \simeq \left(\identmat_d + \sum_i \weight_i \genmat_i + \frac{1}{2} \sum_i \weight_i \genmat_i^2\right)
	\errmat_{\rm eff} \simeq \sum_i \weight_i \left(\identmat_d + \genmat_i + \frac{1}{2}  \genmat_i^2\right)
\end{align}
These error generators lie in a vector space. If, as Campbell required, the origin lies in their convex hull (see Fig.~\ref{fig:vectorspace}), then there exists a choice for the weights, $\weight_i^*$, such that $\sum_i \weight^*=1$ and,
\begin{align}
	\label{eq:cancel}
	&\sum_i \weight_i^* \genmat_i= 0
	% \mathrm{with}\;&\sum_i \weight_i = 1
\end{align}
When this condition is satisfied, we call the resulting operation a \emph{generator-exact} MQG.

If the diamond distance error rates of the component gates are bounded by $\alpha$, then Campbell shows that this first-order cancellation of the error generators can ensure that the diamond distance error rate of the MQG is bounded by $\alpha^2$. Campbell considered error channels that were strictly unitary, so the error generators in Eq.~\eqref{eq:cancel} were Hamiltonians. Hamiltonian generators can have positive or negative coefficients, so it is possible that the origin may lie in their convex hull. If, however, the error map contains stochastic components, then the condition of Eq.~\eqref{eq:cancel} might be impossible to satisfy. Such errors always have strictly positive probabilities, so the origin can \emph{never} lie in their convex hull. This can be rectified by restricting the sum in Eq.~\eqref{eq:cancel} to only the Hamiltonian component of the generators. In the rest of this paper, we will be considering gates subject only to unitary errors, so we will neglect this complication going forward. \note{Use the notation "first-order exact MQG" when "Pauli-exact MQG" to the  }

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{vectorspace.pdf}
  \caption{A target unitary gate can be implemented a number of ways, each associated with a different error generator, $\genmat_i$. These generators lie in a vector space, which we illustrate as two-dimensional here. a) Four error generators. The origin is not contained in their convex hull, so there are no exact MQGs. b) After including an additional control solution, the convex hull grows to contain the origin, and so an exact MQG exists. }
  \label{fig:vectorspace}
\end{figure}

While generator-exact MQGs provide a guaranteed suppression of the diamond distance, their effective error channels are unlikely to be a purely stochastic. Second- and higher-order terms in \eqref{eq:generators_and_effective_error} can easily contribute to lingering coherent errors that impact the efficient simulability of the channel. In order to definitively eliminate these coherent errors, we could instead seek weights, $\weight^P_i$, that annihilate the off-diagonal entries of the unital submatrix of the effective error map. The resulting diagonal error map is Pauli stochastic. For a single qubit gate, it takes the form,
% \begin{equation}
% 	\errmat_{eff} = 
% 	\left(\begin{array}{@{}ccccc@{}}
% 		\cdot & \cdot & \cdot & \cdot \\ 
%   		\cline{1-1}\cline{3-4}
%     	\multicolumn{1}{|c}{0} & \multicolumn{1}{|c}{\cdot} & \multicolumn{1}{|c}{0} & \multicolumn{1}{c|}{0} \\
%     	\cline{2-2}\cline{3-3}		
% 		\multicolumn{1}{|c}{0} & 0 & \multicolumn{1}{|c}{\cdot} & \multicolumn{1}{|c|}{0}\\
% 		\cline{3-3}\cline{4-4}	
% 		\multicolumn{1}{|c}{0} & 0 & 0 & \multicolumn{1}{|c}{\cdot}\\
% 		\cline{1-3}
% 	\end{array} 	
% 	\right).
% \end{equation}
\begin{equation}
	\errmat_{\rm{eff}} = \sum_i \weight_i^P \errmat_i = 
	\left(\begin{array}{@{}ccccc@{}}
		1 & 0 & 0 & 0 \\ 
    	0 &  p_x & 0 & 0 \\
		0 & 0 &  p_y & 0\\
		0 & 0 & 0 &  p_z\\
	\end{array} 	
	\right),
\end{equation}
where $p_x, p_y$ and $p_z$ are the rates of Pauli $X$, $Y$, and $Z$ errors, respectively. We refer to such channels as \emph{Pauli-exact} MQGs, and the same geometric argument we used in constructing generator-exact MQGs is again applicable here. The off-diagonal elements of the error map form a vector space, so we need only check that the origin is in the convex hull of the vectors of off-diagonal elements for each of the component operations, $\errmat_i$.

In the remainder of this paper, we will formalize this process, presenting efficient convex programs to both generate a family of unitary controls and compute the weights required to yeild either type of MQG. We discuss how one might include additional desiderata, including reduced sensitivity to environmental fluctuation as well as limiting the number of native unitaries that participate in the MQG.

>>>>

\newcommand{\actual}{\ensuremath{\tilde{\mathsf{G}}}}
\newcommand{\actmat}{\ensuremath{\tilde{\mathcal{G}}}}
\newcommand{\target}{\ensuremath{{\mathsf{G}}}}
\newcommand{\tarmat}{\ensuremath{{\mathcal{G}}}}
\newcommand{\error}{\ensuremath{{\mathsf{E}}}}
\newcommand{\errmat}{\ensuremath{{\mathcal{E}}}}
\newcommand{\genmat}{\ensuremath{{\mathcal{L}}}}
\newcommand{\vectorize}[1]{\ensuremath{\mathsf{vec}\left(#1\right)}}
\newcommand{\AGI}{\ensuremath{\epsilon_\mathcal{F}}}
\newcommand{\dnorm}{\ensuremath{\epsilon_\diamond}}
\newcommand{\identmat}{\ensuremath{\mathcal{I}}}
\newcommand{\ident}{\ensuremath{\mathsf{I}}}
\newcommand{\0}{\ensuremath{\mathbf{0}}}
\newcommand{\weight}{\ensuremath{w}}

\newcommand{\bunderbrace}[2]{
  \begin{array}[t]{@{}c@{}}
  	#1\\
  	\parbox{\widthof{#1}}{$\scriptscriptstyle#2$}
  \end{array}}

@@

% ==============================================================================
% Section: Introduction
% ==============================================================================
\section{Introduction}
\label{sec:introduction}
\noindent The ultimate impact of a gate error on the performance of a quantum circuit depends strongly on both the magnitude and the nature of the error. Systematic, or \emph{coherent}, errors can arise from poorly calibrated controls or approximate gate compilations that induce repeatable, undesired unitary errors on the state of a quantum information processor (QIP). Errors of this type are correlated in time and may add up constructively or destructively, depending on details of the circuit in which they appear. This can make it difficult to construct tight analytic bounds on circuit performance \cite{Beale2018}, and numerical studies are often limited by the high computational cost of modeling coherent errors. Contrast this against random, or \emph{stochastic}, errors which often result from high-frequency noise in the controls or the environment. Systems with stochastic errors can usually be accurately modeled by defining a rate of various discrete errors in the system, such as a bit flips or phase flips. These errors are significantly easier to simulate on a classical computer, and their impact on quantum circuits is much easier to estimate \cite{Beale2018}.

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{simple_example.pdf}
  \caption{\textbf{Simple example of a mixed quantum gate:} Using optimal control, two implementations of a $Z_\pi$ gate are designed to have equal and opposite coherent errors (if one implementation over-rotates by a small angle $\theta$, then the other \emph{under}-rotates by $\theta$). Each time the gate is used, one of these two implementations is chosen at random. The resulting effective quantum gate is equivalent to a perfect implementation of the gate followed by dephasing with associated probability $\order{\theta^2}$.}
  \label{fig:simple_example}
\end{figure}

Campbell \cite{Campbell2017,1811.08017} and Hastings \cite{Hastings2017} have developed a technique for suppressing coherent noise by replacing deterministic gate implementations with a \emph{mixed quantum gate} (MQG) consisting of a randomly sampled ensemble of implementations. They focus on errors in gate compilation, such as arise from the Solovey-Kitaev algorithm, for which any lingering approximation errors are generally coherent, even if the underlying native gates are perfect. Different approximate gate compilations of the same target unitary will almost certainly result in different unitary errors. So by selecting from these various compilations at random, the resulting quantum channel becomes a mixture of unitaries \cite{DBLP:journals/corr/cs-CC-0012017}, which can have significantly less coherent error than any single compilation on its own. A very simple example of this reduction in coherent error by mixed quantum gates is illustrated in Fig.~\ref{fig:simple_example}.

In this article, we extend the work of Campbell and Hastings to numerically optimized quantum gates and show that the advantages of their approach can be made robust to drift and model uncertainty. We provide an explicit series of convex programs able to efficiently compute MQGs with a variety of useful properties. We demonstrate our results on a superconducting testbed device at Rigetti Computing.  In a simple experiment based on randomized benchmarking circuits, MQGs demonstrate a marked improvement in error rates and a reduced variance in circuit outcome probabilities, consistent with a significant reduction in the coherence of the error \cite{Ball2016}. We further apply our methods in simulation, constructing both single- and two-qubit mixed unitary controls that are robust to drift and uncertainty in the control parameters.

@@

\subsection{Representing Errors in Quantum Gates}
\label{sec:rep_errors}
\noindent In order to implement a desired unitary gate, $\target$, on a quantum device, one generally applies a carefully tuned sequence of classical control fields.  But fluctuations in the environment or imperfections in these controls can cause the state of the qubits to change in a way that is different from what was intended, \emph{i.e.}, there are errors in the gate. If the device is fairly stable with time\kcy{cite drift paper} and context\cite{Rudinger2019}, then we can accurately model the gate action using a completely positive, trace-preserving (CPTP) map, $\actual$ acting on the Hilbert space of the target qubits. This map can always be written as $\actual = \error\circ\target$, where $\error = \actual\circ\target^{-1}$ is the \emph{error map}, which is itself CPTP because $\target$ is unitary.

CPTP maps possess a number of useful representations, including Kraus operators\cite{1983}, Choi matrices\cite{Choi1975}, and Jamiolkowsi states\cite{yczkowski2004}. But for the purposes of this article, the \emph{process matrix} representation will be particularly convenient\cite{OBrien2004}, and we shall denote the process matrix associated with a given CPTP map $\target$ with its corresponding calligraphic character, $\tarmat$. For a $d$-dimensional quantum state, the process matrix is a $d^2\times d^2$ matrix. A key feature of process matrices is that they are composable and act through the usual matrix multiplication on the vectorized quantum state:
\begin{align}
	\vectorize{\actual(\rho)}
		&= \actmat\cdot\vectorize{\rho} \\
		&= \errmat\cdot\tarmat\cdot\vectorize{\rho}
\end{align}
The $\mathsf{vec}$ operation is typically performed using a basis of matrix units, for which $\vectorize{\rho}$ would be the column vector obtained by stacking the columns of $\rho$. In this work, however, we shall use a basis of Pauli matrices, defining $P = \left\{I, \sigma_x, \sigma_y, \sigma_z\right\}^{\otimes n}$ as the collection of all $4^n$ $n$-qubit Pauli operators (including the identity).  In this basis,
\begin{equation}
  \vectorize{\rho}_i = \expect{P_i} = \tr\!\left(P_i\; \rho \right),
\end{equation}
and
\begin{equation}
	(\actmat)_{ij} = \tr\!\left(P_i \;\actual\!\left(P_j\right) \right).
\end{equation}
Process matrices written in this Pauli basis are often referred to as \emph{Pauli transfer matrices} \cite{Chow2012}.  Error maps represented in this basis take the particularly nice form,
\begin{equation}\label{eq:process_matrix}
\errmat =
	\left(\begin{array}{c|cccc}
		1 &  & \vec{0}^T & \\ 
		\hline & &  &  \\
		\vec{m} &  & R &  \\
		 &  &  & 
	\end{array} 	
	\right)
\end{equation}
The top row of all trace-preserving (TP) error maps is fixed to $\{1,0,0,0,\cdots\}$ and the remainder of the first column, $\vec{m}$, describes any deviations from unitality, as might arise from amplitude damping \cite{preskill1997lecture}. If the error map is unitary, then the error is called \textit{coherent} and the unital submatrix $R$ is a rotation matrix.  Importantly, if $R$ is diagonal, then the error is Pauli stochastic, with each diagonal entry corresponding to the probability that its associated Pauli error occurs in each application of the gate. \kcy{We might need more here defining coherent noise. There is a lot of literature on this, most of it inconsistent. Try Joel's Estimating the Coherence of Noise (terrible title) and Gottesman's paper on Maximally Sensitive States.}

In what follows it will be useful to define the \emph{error generator}, $\genmat$, associated with a faulty gate: 
\begin{align}
	\errmat 
		&= \exp\left(\genmat\right) \\
	\label{eq:generator}
		&= \identmat_d + \genmat + \frac{1}{2}\genmat^2 + \order{\genmat^3}.
\end{align}
If an implemented gate is relatively close to its target, then the error generator will be small under any of the usual matrix norms, and the Taylor expansion above may reliably be truncated at first or second order.

\subsection{Mixed Quantum Gates}
\label{sec:mqg}
\noindent Suppose that we have access to an ensemble of distinct implementations, 
$\{\actual_i\}$, of a target gate, $\target$.
Each time the gate is to be applied to the system, we randomly select an implementation from this ensemble such that the probability of drawing $\actual_i$ is $\weight_i$ (and we ensure that $\sum_i \weight_i=1$). This procedure is operationally indistinguishable from always applying the effective channel, 
\begin{align}
	\actual_{\rm eff} &= \sum_i \weight_i \actual_i = \left(\sum_i \weight_i \error_i \right) \circ\target
\end{align}
We call such randomized quantum operations \emph{mixed quantum gates} or MQGs. Error metrics for these MQGs can be computed in terms of their effective error map, 
\begin{equation}
	\label{eq:effective_error}
	\error_{\rm eff} = \sum_i \weight_i \error_i.
\end{equation}
Two important error metrics are the average gate infidelity (AGI), $\AGI$, \kcy{We need a citation for this, and check if it's right! Gives weird answer if E = I. I feel like it should be $(d + \tr(\errmat))/(d+d^2)$. Maybe cite Nielsen's paper, or Horodecki**3.} and the diamond distance to the target (or simply, the \emph{diamond distance}), $\dnorm$ \cite{watrous2018theory}, defined as:
\begin{align}
	\label{eq:def_agi}
	&\AGI(\error) = \frac{d^2 - \tr{\errmat}}{d^2 + d},\\
	\label{eq:def_diamond}
	&\dnorm \left(\error\right)
		= \frac{1}{2} \sup_\rho \vert \vert (\ident_d\otimes \ident_d)(\rho) 
										  - (\error \otimes \ident_d)(\rho) \vert\vert_1,
\end{align}
where $d=2^n$ and $\ident_d$ is the $d$-dimensional identity operator.  In Eq.~\eqref{eq:def_agi}, we have written the AGI for an error map, $\error$, in terms of its associated Pauli transfer matrix, $\errmat$. If the error channel is purely stochastic, then $\dnorm(\error) = \AGI(\error)$, but if the error channel has a unitary component, then the diamond distance will generically be larger than the average gate infidelity \cite{1511.00727}. The diamond distance is subaddative \cite{watrous2018theory} under gate composition, so is particularly useful for constructing bounds on quantum circuit performance: the total variation distance between the outcome probabilities of a faulty and perfect quantum circuit is less than or equal to the sum of the diamond distances for gates that compose the circuit \cite{aharonov1998quantum}.

@@

Campbell \cite{Campbell2017} and Hastings \cite{1612.01011} independently considered these mixed channels using gates constructed with the Solovey-Kitaev algorithm, for which many approximate gate compilations are possible. Campbell showed that, if the error generators of some ensemble of gate compilations form a convex set containing the origin, then one can construct a MQG with quadratically suppressed diamond distance to the target. Explicitly, the weights are chosen such that the error generator is canceled to first order. Using (\ref{eq:effective_error}) and (\ref{eq:generator}), the effective error map for a MQG in terms of the component error generators is,
\begin{align}
	\label{eq:generators_and_effective_error}
	% \errmat_{\rm eff} \simeq \left(\identmat_d + \sum_i \weight_i \genmat_i + \frac{1}{2} \sum_i \weight_i \genmat_i^2\right)
	\errmat_{\rm eff} \simeq \sum_i \weight_i \left(\identmat_d + \genmat_i + \frac{1}{2}  \genmat_i^2\right)
\end{align}
These error generators are linear operators and so are elements of a vector space. If, as Campbell required, the origin lies in their convex hull (see Fig.~\ref{fig:vectorspace}), then there exists a choice for the weights, $\weight_i^*$, such that $\sum_i \weight^*=1$ and,
\begin{align}
	\label{eq:cancel}
	&\sum_i \weight_i^* \genmat_i= 0
	% \mathrm{with}\;&\sum_i \weight_i = 1
\end{align}
When this condition is satisfied, we call the resulting operation a \emph{generator-exact} MQG.

If the diamond distance error rates of the component gates are bounded by $\alpha$, then Campbell shows that this first-order cancellation of the error generators can ensure that the diamond distance error rate of the MQG is bounded by $\alpha^2$. For gate compilation errors, the error channels are dominated by unitary approximation error, so the error generators are Hamiltonians. Hamiltonian generators can have positive or negative coefficients, so it is possible that the origin may lie in their convex hull. If, however, the error map contains stochastic components, then the condition of Eq.~\eqref{eq:cancel} might be impossible to satisfy. Such errors always have strictly positive probabilities, so the origin can \emph{never} lie in their convex hull. This can be rectified by restricting the sum in Eq.~\eqref{eq:cancel} to only the Hamiltonian component of the generators, or by replacing the exact condition with a minimization, as we discuss in Sec.~\ref{sec:mixed_unitary_processes}.

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{vectorspace.pdf}
  \caption{\textbf{Geometric condition for existence:} A target unitary gate can be implemented a number of ways, each associated with a different error generator, $\genmat_i$. These generators lie in a vector space, which we illustrate as two-dimensional here. a) Four error generators. The origin is not contained in their convex hull, so there are no generator-exact MQGs. b) After including an additional control solution, the convex hull grows to contain the origin, and so a generator-exact MQG exists. }
  \label{fig:vectorspace}
\end{figure}

While generator-exact MQGs can provide a guaranteed suppression of the diamond distance, their effective error channels are unlikely to be a purely stochastic. Second- and higher-order terms in \eqref{eq:generators_and_effective_error} can easily contribute to lingering coherent errors that impact the efficient simulability of the channel. In order to definitively eliminate these coherent errors, we could instead seek weights, $\weight^P_i$, that annihilate the off-diagonal entries of the effective error map. The resulting diagonal error map will be Pauli stochastic. For a single qubit gate, it takes the form,
% \begin{equation}
% 	\errmat_{eff} = 
% 	\left(\begin{array}{@{}ccccc@{}}
% 		\cdot & \cdot & \cdot & \cdot \\ 
%   		\cline{1-1}\cline{3-4}
%     	\multicolumn{1}{|c}{0} & \multicolumn{1}{|c}{\cdot} & \multicolumn{1}{|c}{0} & \multicolumn{1}{c|}{0} \\
%     	\cline{2-2}\cline{3-3}		
% 		\multicolumn{1}{|c}{0} & 0 & \multicolumn{1}{|c}{\cdot} & \multicolumn{1}{|c|}{0}\\
% 		\cline{3-3}\cline{4-4}	
% 		\multicolumn{1}{|c}{0} & 0 & 0 & \multicolumn{1}{|c}{\cdot}\\
% 		\cline{1-3}
% 	\end{array} 	
% 	\right).
% \end{equation}
\begin{equation}
	\errmat_{\rm{eff}} = \sum_i \weight_i^P \errmat_i = 
	\left(\begin{array}{@{}ccccc@{}}
		1 & 0 & 0 & 0 \\ 
    	0 &  p_x & 0 & 0 \\
		0 & 0 &  p_y & 0\\
		0 & 0 & 0 &  p_z\\
	\end{array} 	
	\right),
\end{equation}
where $p_x, p_y$ and $p_z$ are the rates of Pauli $X$, $Y$, and $Z$ errors, respectively. We refer to such channels as \emph{Pauli-exact} MQGs, and the same geometric argument we used in constructing generator-exact MQGs is again applicable here. The off-diagonal elements of the error map form a vector space, so we need only check that the origin is in the convex hull of the vectors of off-diagonal elements for each of the component operations, $\errmat_i$. If the error map contains non-unital components or correlated Pauli stochastic errors (such as $X+Y$), then this strict geometric condition will not hold and no Pauli-exact MQG will exist. However, MQGs that approximate the Pauli-exact condition can be found via the convex programing techniques discussed in Sec.~\ref{sub:off_diagonals}.

% In the remainder of this paper, we will formalize this process, laying out efficient convex programs to both generate a family of unitary controls \kcy{do we do this?} and compute the weights required to yield either type of MQG. We discuss how one might include additional desiderata, including reduced sensitivity to environmental fluctuation as well as limiting the number of native unitaries that participate in the MQG.

% ==============================================================================
% Section: Constructing Useful mixed quantum gates
% ==============================================================================
\section{Constructing mixed quantum gates}
\label{sec:mixed_unitary_processes}
\noindent In this section, we present a methodology for constructing mixed quantum channels, formalizing the intuitive approach discussed above. As mentioned, our method requires two steps. The first is a control synthesis step, in which we construct an ensemble of gate implementations. Campbell and Hastings draw their ensembles from various different gate compilations, but for this work we use utilize quantum optimal control theory. Many standard control-generation algorithms, such as GRAPE \cite{Khaneja2005}, take a random initial guess for the control and iteratively improve it to yield a gate that well-approximates a target. By seeding such algorithms with many initial guesses, one can quickly construct a large ensemble of approximate quantum gates, each with a different error channel. We discuss this approach in some detail in Sec.~\ref{sec:numerical_results}. \kcy{Why should the gates we generate have any error at all? They shouldn't if we know the model. Discuss this here.}

% ==============================================================================
% Section: Generators
% ==============================================================================
\subsection{Convex Programs for Constructing Mixed Quantum Gates}
\subsubsection{Generator-exact MQGs} % (fold)
\label{sub:first_order_generators}
\noindent A compelling reason to construct an MQG is to improve the worst-case performance of a quantum gate. The diamond distance bounds this worst-case performance, and so one might assume minimizing it would be a natural optimization target:
\begin{equation}\label{eq:ddmin}
  \begin{split}
    &\underset{\weight_i\geq0, \sum_i\weight_i=1}{\textbf{minimize}: } \hspace{.5cm} \dnorm\left(\sum_i\error_i \weight_i\right)
  \end{split}
\end{equation}
The diamond distance, however, is a non-linear function that in general requires its own convex optimization to compute\cite{watrous2018theory}. This can dramatically slow down iterative optimizers, and so directly minimizing the diamond distance is a computationally burdensome optimization target. However, for quantum computing purposes the error rates are typically quite small, so we can consider a linearized problem, minimizing the diamond distance at first-order in the effective error generator. As discussed around Eq.~\eqref{eq:cancel}, if the origin lies in in the convex hull of the generators, $\0 \in \mathsf{conv}(\{\genmat_i\})$, then we can construct a generator-exact MQG. In \cite{Campbell2017} Campbell presents an iterative algorithm that, given an oracle able to produce approximate unitary operations, will continually generate controls until the geometric constraint is satisfied and then identify the optimal weighting.

As discussed in the previous section, the geometric condition illustrated in Fig.~\ref{fig:vectorspace} may not be satisfied, due perhaps to stochastic errors in the generators or possible a situation in which one has access only to $m$ distinct, precomputed gate implementations. In this case, we may simply wish to to identify the MQG with minimum error. A heuristic that is efficiently computable is to minimize the Frobenius norm of the first-order effective error generator:
\begin{equation}\label{eq:frobmin}
  \begin{split}
    &\underset{\weight_i\geq0, \sum_i\weight_i=1}{\textbf{minimize}: } \hspace{.5cm}  \left\vert\left\vert\sum_i \weight_i \genmat_i \right\vert\right\vert_F
  \end{split}
\end{equation}
This optimization problem is equivalent to:
\begin{equation}\label{eq:minimization}
  \begin{split}
    &\underset{\weight_i\geq0, |\weight|_1=1}{\textbf{minimize}: } ||\mathbf{L}\cdot\vec\weight||_2\\
  \end{split}
\end{equation}
Where $\mathbf{L}=\left( \mathsf{col}(\genmat_1) \; \mathsf{col}(\genmat_2) \; \cdots \;  \right)$ is the $d^2\times m$-dimensional matrix of column-stacked error generators and $||\cdot||_2$ is the $\ell_2$ norm.
